Supplementary Material \\\large{Shifts in water supply and demand drive land cover change across Chile}

### 2024-06-25 article/article_nature/supplementary_material.qmd
actualizar repositorio0+

---
title: "Supplementary Material"
subtitle: "The effects of drought on land cover change and vegetation productivity in continental Chile"
author:
  - name: Francisco Zambrano
    email: francisco.zambrano@umayor.cl
    affiliations: 
        - id: umayor
          name: Hémera Centro de Observación de la Tierra, Facultad de Ciencias, Escuela de Ingeniería en Medio Ambiente y Sustentabilidad, Universidad Mayor
          city: Santiago
          country: 7500994, Chile.
        - id: odes
          name: Observatorio de Sequía para la Agricultura y la Biodiversidad de Chile (ODES), Universidad Mayor
          city: Santiago
          country: 7500994, Chile.
    attributes:
        corresponding: true
    #note: This is the first author footnote.
  - name: Anton Vrieling
    #email: a.vrieling@utwente.nl
    affiliations:
        - id: itc
          name: Faculty of Geo-Information Science and Earth, University of Twente
          city:  Enschede
          country: The Netherlands.
  - name: Francisco Meza
    affiliations:
        - id: agrouc
          name: Facultad de Agronomía y Sistemas Naturales, Pontificia Universidad Católica de Chile. 
          city: Santiago
          country: Chile.
        - id: instuc
          name: Instituto para el Desarrollo Sustentable. Pontificia Universidad Católica de Chile
          city: Santiago
          country: Chile.
        - id: cambioglobal
          name: Centro Interdisciplinario de Cambio Global, Pontificia Universidad Católica de Chile
          city: Santiago
          country: Chile.
  - name: Iongel Duran-Llacer
    affiliations:
        - id: umayor-io
          name: Hémera Centro de Observación de la Tierra, Facultad de Ciencias, Universidad Mayor,
          city: Santiago
          country: 7500994, Chile.
  - name: Francisco Fernández
    #email: francisco.fernandez@uss.cl
    affiliations:
        - id: uss
          name: Center of Economics for Sustainable Development (CEDES), Faculty of Economics and Government, Universidad San Sebastian
          city: Santiago
          country: Chile.
        - id: capes
          name: Center of Applied Ecology and Sustainability (CAPES)
          city: Santiago
          country: Chile.
  - name: Alejandro Venegas-González
    affiliations:
        - id: ica3
          name: Instituto de Ciencias Agroalimentarias, Animales y Ambientales (ICA3), Universidad de O'Higgins
          city: San Fernando
          country: Chile.
  - name: Nicolas Raab
    affiliations:
        - id: agrouc
  - name: Dylan Craven
    affiliations:
        - id: gema
          department: GEMA Center for Genomics, Ecology & Environment, Universidad Mayor, Camino La Pirámide Huechuraba 5750
          city: Santiago
          country: Chile.
        - id: do
          name: Data Observatory Foundation
          city: Santiago
          country: Chile.
date: "`r Sys.Date()`"
include-before-body:
  - text: |
      \renewcommand\thesection{S\arabic{section}}
      \renewcommand\thefigure{S\arabic{figure}}
      \renewcommand\thetable{S\arabic{table}}
bibliography: references.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      layout: onecolumn
      model: 3p
      cite-style: authoryear
      natbiboptions: numbers
    include-in-header: 
      text: \usepackage{lineno}\linenumbers
            \usepackage{multirow}
            \usepackage{lscape}
            \newcommand{\blandscape}{\begin{landscape}}
            \newcommand{\elandscape}{\end{landscape}}
editor: source
---

@@

@@

```{=latex}
\begin{table}[!ht]
\caption{Landcover clases from IGBP MODIS and the corresponding macroclasses.}
\label{tab-landIGBPmacro}
\centering
\includegraphics[width = .5\textwidth]{../output/figs/tabla_landcover_macroclass.png}
\end{table}
```

@@

```{r fig-RF_importance}
#| echo: false
#| layout-nrow: 2
#| fig-pos: "!ht"
#| fig-env: "figure*"
#| out-width: 100%
#| fig-cap: "Error bar for the variables importance obtained from the 10 resamples folded by the Random Forest used to model land cover trend using as predictors the trend in drought indices of water supply, water demand, soil moisture, and vegetation productivity."
#| fig-subcap: 
#| - "Forest"
#| - "Cropland"
#| - "Grassland" 
#| - "Savanna"
#| - "Shrubland"
#| - "Barren Land"

knitr::include_graphics("../output/figs/fig_errorbar_resample_random_forest_trends_Forest.png")
knitr::include_graphics("../output/figs/fig_errorbar_resample_random_forest_trends_Cropland.png")
knitr::include_graphics("../output/figs/fig_errorbar_resample_random_forest_trends_Grassland.png")
knitr::include_graphics("../output/figs/fig_errorbar_resample_random_forest_trends_Savanna.png")
knitr::include_graphics("../output/figs/fig_errorbar_resample_random_forest_trends_Shrubland.png")
knitr::include_graphics("../output/figs/fig_errorbar_resample_random_forest_trends_Barren_land.png")
```

@@

```{r fig-hetmaptrendzcNDVI}
#| echo: false
#| out-width: 75%
#| fig-pos: '!ht'
#| fig-cap: "Heatmap of trends in zcNDVI for 2000 to 2023 per macrozone and landcover macroclass."
knitr::include_graphics('../output/figs/heatmap_trends_zcNDVI_macro_landcover.png')
```

@@

```{r fig-r2_npp_zcNDVI}
#| echo: false
#| out-width: 75%
#| fig-pos: '!ht'
#| fig-cap: "Spatial variation of the r-squared values obtained from the yearly correlation of zcNDVI of 1, 3, 6, and 12 months with the net primary productivity (NPP) for continental Chile."
knitr::include_graphics('../output/figs/map_r2_NPP_vs_zcNDVI.png')
```

```{r fig-hetmap_npp_zcndvi}
#| echo: false
#| out-width: 50%
#| fig-pos: '!ht'
#| fig-cap: "A heatmap showing the r-squared values obtained from the yearly correlation of zcNDVI of 1, 3, 6, and 12 months with the net primary productivity (NPP) for continental Chile."
knitr::include_graphics('../output/figs/heatmap_r2_npp_vs_zcndvi_per_landcover.png')
```

# Relationship of drought indices and vegetation productivity

```{r}
#| echo: false
#| out-width: 100%
#| fig-pos: '!ht'
#| fig-cap: "Summary per land cover and study region regarding the correlation between zcNDVI with the drought indices EDDI, SPI, SPEI, and SSI for time scales of 1, 3, 6, 12, 24, and 36. The numbers in each cell indicate the time scale that reached the maximum correlation for the land cover and macrozone, and the color indicates the strength of the r-squared obtained with the index and the time scale. Cells without values indicate that the land cover type was not significant in that macrozone."
knitr::include_graphics('../output/figs/tabla_r_cor_macro_indice.png')
```

>>>>

---
title: "Supplementary Material"
subtitle: "Shifts in water supply and demand drive land cover change across Chile"
author:
  - name: Francisco Zambrano
    email: francisco.zambrano@umayor.cl
    affiliations: 
        - id: umayor
          name: Hémera Centro de Observación de la Tierra, Facultad de Ciencias, Escuela de Ingeniería en Medio Ambiente y Sustentabilidad, Universidad Mayor
          city: Santiago
          country: 7500994, Chile.
        - id: odes
          name: Observatorio de Sequía para la Agricultura y la Biodiversidad de Chile (ODES), Universidad Mayor
          city: Santiago
          country: 7500994, Chile.
    attributes:
        corresponding: true
    #note: This is the first author footnote.
  - name: Anton Vrieling
    #email: a.vrieling@utwente.nl
    affiliations:
        - id: itc
          name: Faculty of Geo-Information Science and Earth, University of Twente
          city:  Enschede
          country: The Netherlands.
  - name: Francisco Meza
    affiliations:
        - id: agrouc
          name: Facultad de Agronomía y Sistemas Naturales, Pontificia Universidad Católica de Chile. 
          city: Santiago
          country: Chile.
        - id: instuc
          name: Instituto para el Desarrollo Sustentable. Pontificia Universidad Católica de Chile
          city: Santiago
          country: Chile.
        - id: cambioglobal
          name: Centro Interdisciplinario de Cambio Global, Pontificia Universidad Católica de Chile
          city: Santiago
          country: Chile.
  - name: Iongel Duran-Llacer
    affiliations:
        - id: umayor-io
          name: Hémera Centro de Observación de la Tierra, Facultad de Ciencias, Universidad Mayor,
          city: Santiago
          country: 7500994, Chile.
  - name: Francisco Fernández
    #email: francisco.fernandez@uss.cl
    affiliations:
        - id: uss
          name: Center of Economics for Sustainable Development (CEDES), Faculty of Economics and Government, Universidad San Sebastian
          city: Santiago
          country: Chile.
        - id: capes
          name: Center of Applied Ecology and Sustainability (CAPES)
          city: Santiago
          country: Chile.
  - name: Alejandro Venegas-González
    affiliations:
        - id: ica3
          name: Instituto de Ciencias Agroalimentarias, Animales y Ambientales (ICA3), Universidad de O'Higgins
          city: San Fernando
          country: Chile.
  - name: Nicolas Raab
    affiliations:
        - id: agrouc
  - name: Dylan Craven
    affiliations:
        - id: gema
          department: GEMA Center for Genomics, Ecology & Environment, Universidad Mayor, Camino La Pirámide Huechuraba 5750
          city: Santiago
          country: Chile.
        - id: do
          name: Data Observatory Foundation
          city: Santiago
          country: Chile.
date: "`r Sys.Date()`"
include-before-body:
  - text: |
      \renewcommand\thesection{S\arabic{section}}
      \renewcommand\thefigure{S\arabic{figure}}
      \renewcommand\thetable{S\arabic{table}}
bibliography: references.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      layout: onecolumn
      model: 3p
      cite-style: authoryear
      natbiboptions: numbers
    include-in-header: 
      text: \usepackage{lineno}\linenumbers
            \usepackage{multirow}
            \usepackage{lscape}
            \newcommand{\blandscape}{\begin{landscape}}
            \newcommand{\elandscape}{\end{landscape}}
editor: source
---

@@

# Results

```{r}
#| echo: false
#| out-width: 100%
#| tbl-pos: '!ht'
#| tbl-cap: "Summary per land cover and study region regarding the correlation between zcNDVI with the drought indices EDDI, SPI, SPEI, and SSI for time scales of 1, 3, 6, 12, 24, and 36. The numbers in each cell indicate the time scale that reached the maximum correlation for the land cover and macrozone, and the color indicates the strength of the r-squared obtained with the index and the time scale. Cells without values indicate that the land cover type was not significant in that macrozone."
knitr::include_graphics('figs/tabla_r_cor_macro_indice.png')
```

@@

```{=latex}
\begin{table}[!ht]
\caption{Landcover clases from IGBP MODIS and the corresponding macroclasses.}
\label{tab-landIGBPmacro}
\centering
\includegraphics[width = .5\textwidth]{figs/tabla_landcover_macroclass.png}
\end{table}
```

@@

```{r fig-RF_importance}
#| echo: false
#| layout-nrow: 2
#| fig-pos: "!ht"
#| fig-env: "figure*"
#| out-width: 100%
#| fig-cap: "Error bar for the variables importance obtained from the 10 resamples folded by the random forest used to model land cover trend using as predictors the trend in drought indices of water supply, water demand, soil moisture, and vegetation productivity."
#| fig-subcap: 
#| - "Forest"
#| - "Cropland"
#| - "Grassland" 
#| - "Savanna"
#| - "Shrubland"
#| - "Barren Land"
knitr::include_graphics("figs/fig_errorbar_resample_random_forest_trends_Forest_norte grande.png")
knitr::include_graphics("figs/fig_errorbar_resample_random_forest_trends_Cropland_norte grande.png")
knitr::include_graphics("figs/fig_errorbar_resample_random_forest_trends_Grassland_norte grande.png")
knitr::include_graphics("figs/fig_errorbar_resample_random_forest_trends_Savanna_norte grande.png")
knitr::include_graphics("figs/fig_errorbar_resample_random_forest_trends_Shrubland_norte grande.png")
knitr::include_graphics("figs/fig_errorbar_resample_random_forest_trends_Barren_land_norte grande.png")
```

@@

```{r fig-hetmaptrendzcNDVI}
#| echo: false
#| out-width: 75%
#| fig-pos: '!ht'
#| fig-cap: "Heatmap of trends in zcNDVI for 2000 to 2023 per macrozone and landcover macroclass."
knitr::include_graphics('figs/heatmap_trends_zcNDVI_macro_landcover.png')
```

@@

```{r fig-r2_npp_zcNDVI}
#| echo: false
#| out-width: 75%
#| fig-pos: '!ht'
#| fig-cap: "Spatial variation of the r-squared values obtained from the yearly correlation of zcNDVI of 1, 3, 6, and 12 months with the net primary productivity (NPP) for continental Chile."
knitr::include_graphics('figs/map_r2_NPP_vs_zcNDVI.png')
```

```{r fig-hetmap_npp_zcndvi}
#| echo: false
#| out-width: 50%
#| fig-pos: '!ht'
#| fig-cap: "A heatmap showing the r-squared values obtained from the yearly correlation of zcNDVI of 1, 3, 6, and 12 months with the net primary productivity (NPP) for continental Chile."
knitr::include_graphics('figs/heatmap_r2_npp_vs_zcndvi_per_landcover.png')
```

### 2024-06-25 article/article_nature/supplementary_material.tex
actualizar repositorio0+

\begin{frontmatter}
\title{Supplementary Material \\\large{The effects of drought on land
cover change and vegetation productivity in continental Chile} }
\author[1,2]{Francisco Zambrano%
\corref{cor1}%
}
 \ead{francisco.zambrano@umayor.cl} 
\author[3]{Anton Vrieling%
%
}

@@

@@

\subsection{Results}\label{results}

@@

@@

\begin{table}[!ht]
\caption{Landcover clases from IGBP MODIS and the corresponding macroclasses.}
\label{tab-landIGBPmacro}
\centering
\includegraphics[width = .5\textwidth]{../output/figs/tabla_landcover_macroclass.png}
\end{table}

@@

\includegraphics[width=1\textwidth,height=\textheight]{../output/figs/fig_errorbar_resample_random_forest_trends_Forest.png}

@@

\includegraphics[width=1\textwidth,height=\textheight]{../output/figs/fig_errorbar_resample_random_forest_trends_Cropland.png}

@@

\includegraphics[width=1\textwidth,height=\textheight]{../output/figs/fig_errorbar_resample_random_forest_trends_Grassland.png}

@@

\includegraphics[width=1\textwidth,height=\textheight]{../output/figs/fig_errorbar_resample_random_forest_trends_Savanna.png}

@@

\includegraphics[width=1\textwidth,height=\textheight]{../output/figs/fig_errorbar_resample_random_forest_trends_Shrubland.png}

@@

\includegraphics[width=1\textwidth,height=\textheight]{../output/figs/fig_errorbar_resample_random_forest_trends_Barren_land.png}

@@

\caption{\label{fig-RF_importance}Error bar for the variables importance
obtained from the 10 resamples folded by the Random Forest used to model
land cover trend using as predictors the trend in drought indices of
water supply, water demand, soil moisture, and vegetation productivity.}

@@

\includegraphics[width=0.75\textwidth,height=\textheight]{../output/figs/heatmap_trends_zcNDVI_macro_landcover.png}

@@

\includegraphics[width=0.75\textwidth,height=\textheight]{../output/figs/map_r2_NPP_vs_zcNDVI.png}

@@

\includegraphics[width=0.5\textwidth,height=\textheight]{../output/figs/heatmap_r2_npp_vs_zcndvi_per_landcover.png}

@@

\section{Relationship of drought indices and vegetation
productivity}\label{relationship-of-drought-indices-and-vegetation-productivity}

\begin{figure}[H]

{\centering \includegraphics[width=1\textwidth,height=\textheight]{../output/figs/tabla_r_cor_macro_indice.png}

\caption{Summary per land cover and study region regarding the
correlation between zcNDVI with the drought indices EDDI, SPI, SPEI, and
SSI for time scales of 1, 3, 6, 12, 24, and 36. The numbers in each cell
indicate the time scale that reached the maximum correlation for the
land cover and macrozone, and the color indicates the strength of the
r-squared obtained with the index and the time scale. Cells without
values indicate that the land cover type was not significant in that
macrozone.}

>>>>

\begin{frontmatter}
\title{Supplementary Material \\\large{Shifts in water supply and demand
drive land cover change across Chile} }
\author[1,2]{Francisco Zambrano%
\corref{cor1}%
}
 \ead{francisco.zambrano@umayor.cl} 
\author[3]{Anton Vrieling%
%
}

@@

\section{Results}\label{results}

\includegraphics[width=1\textwidth,height=\textheight]{figs/tabla_r_cor_macro_indice.png}

@@

@@

\subsection{Results}\label{results-2}

@@

\begin{table}[!ht]
\caption{Landcover clases from IGBP MODIS and the corresponding macroclasses.}
\label{tab-landIGBPmacro}
\centering
\includegraphics[width = .5\textwidth]{figs/tabla_landcover_macroclass.png}
\end{table}

@@

\includegraphics[width=1\textwidth,height=\textheight]{figs/fig_errorbar_resample_random_forest_trends_Forest_norte grande.png}

@@

\includegraphics[width=1\textwidth,height=\textheight]{figs/fig_errorbar_resample_random_forest_trends_Cropland_norte grande.png}

@@

\includegraphics[width=1\textwidth,height=\textheight]{figs/fig_errorbar_resample_random_forest_trends_Grassland_norte grande.png}

@@

\includegraphics[width=1\textwidth,height=\textheight]{figs/fig_errorbar_resample_random_forest_trends_Savanna_norte grande.png}

@@

\includegraphics[width=1\textwidth,height=\textheight]{figs/fig_errorbar_resample_random_forest_trends_Shrubland_norte grande.png}

@@

\includegraphics[width=1\textwidth,height=\textheight]{figs/fig_errorbar_resample_random_forest_trends_Barren_land_norte grande.png}

@@

\caption{\label{fig-RF_importance}Error bar for the variables importance
obtained from the 10 resamples folded by the random forest used to model
land cover trend using as predictors the trend in drought indices of
water supply, water demand, soil moisture, and vegetation productivity.}

@@

\includegraphics[width=0.75\textwidth,height=\textheight]{figs/heatmap_trends_zcNDVI_macro_landcover.png}

@@

\includegraphics[width=0.75\textwidth,height=\textheight]{figs/map_r2_NPP_vs_zcNDVI.png}

@@

\includegraphics[width=0.5\textwidth,height=\textheight]{figs/heatmap_r2_npp_vs_zcndvi_per_landcover.png}

@@
